---
title: "Tourism Economics of Nepal"
author:
- affiliation: EMBA, Ace Institute of Management
  name: Ashish Singh
date: "October 3, 2016"
output:
  pdf_document:
    citation_package: natbib
    fig_caption: yes
    latex_engine: pdflatex
    template: svm-latex-ms.tex
  html_document: default
  word_document: default
fontsize: 11pt
geometry: margin=1in
fontfamily: mathpazo
subtitle: Analysis & forecasting
abstract: Trend analysis and forecasting of various tourism related economic indicators
  of Nepal. The time series data used ranges from the year 2002 to 2015 and forecasting
  is done for the subsequent three years. Variables used are Total Foreign Exchange
  Reserves, Number of Tourist Arrival, Gross Foreign Earnings from Tourism, Average
  Length of Stay and Travel & Tourism Direct Contribution to Employment.
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Chapter 1: Introduction

## Impact of Tourism
Generating US$7.6 trillion (10% of global GDP) and 277 million jobs (1 in 11 jobs) for the
global economy in 2014, Travel & Tourism is an important economic activity in most countries around the world. As well as its direct economic
impact, the sector has significant indirect and induced impacts. Recent years have seen Travel & Tourism growing at a faster rate than both the
wider economy and other significant sectors such as automotive, financial services and health care. Last year
was no exception. International tourist arrivals also surged, reaching nearly 1.14billion and visitor spending
more than matched that growth. Visitors from emerging economies now represent a 46% share of these
international arrivals (up from 38% in 2000), proving the growth and increased opportunities for travel from
those in these new markets (Turner, 2015, pg. 2).

## Toursim in Nepal
Nepal Tourism Policy 2009 identifies tourism sector as an important
vehicle for economic and social development. Following directions set by the
policy effort has been put for development and expansion of tourism activities,
quality improvement of tourism services, increasing revenue and expansion of
employment opportunities to improve the living standard of Nepalese people.
Vision 2020 of tourism envisions increasing tourist arrival to two million and
tourism related employment to one million (Ministry of Culture, Tourism & Civil Aviation, 2016). 



```{r echo = FALSE}
library(ggplot2)
library(reshape2)
library(knitr)
library(formattable)
options("scipen"=100, "digits"=2)

nepal_tourism <- read.table(file = "/mnt/765E305F5E3019F7/Documents and Settings/Ashish Singh/My Documents/My Docs/Ace IoM/Semester II/Data Modeling II/nepal_economic_survery.csv", header = TRUE, sep=",", check.names = FALSE)
#rownames(nepal_tourism) <- nepal_tourism[,1]
#nepal_tourism <- nepal_tourism[,-1]

precise_tourism <- cbind(nepal_tourism[,1], sprintf("%.2f", nepal_tourism[, 2]), nepal_tourism[,3], sprintf("%.3f", nepal_tourism[, 4]), nepal_tourism[, 5], nepal_tourism[, 6])

# Formatting precision for display
nepal_tourism$`Total Foreign Exchange Reserves` = formattable(nepal_tourism[, 2], digits=2, format="f")
nepal_tourism$`Gross Foreign Exchange Earnings From Tourism` = formattable(nepal_tourism[, 4] , digits=3, format="f")

```



```{r fig.width=5, fig.height=3, fig.cap="Even with sudden decline of number of tourists after 2015 earthquake, we can see that there was a steady growth in the number of employnent directly contributed by the tourism industry.", echo=FALSE}
#tsforex <- ts(nepal_tourism[,2], frequency = 1, start = 2002)
#plot.ts(tsforex)
ggplot(nepal_tourism, aes(x=Year)) + geom_line(aes(y=`Tourist Arrival`, linetype="Tourist Arrival")) + geom_smooth(aes(y=`Travel & Tourism Direct Contribution to Employment`, linetype="Direct Contribution to Employment"), se=FALSE, color="#444444") + xlab("Year") + ylab("Number (in Thousands)") + theme_light()  + guides(linetype=guide_legend(title="")) + annotate(geom="text", x=2014.8, y=700000, label="Decline ->", color="red", angle = -70) + theme(legend.position="top", text = element_text(size=10)) + scale_y_continuous(labels = function(x)x/1000)
```


Trend of tourist arrival was in increased order up to 2012 which
declined thereafter. After the 7.8 magnitude earthquake that rocked Nepal, we saw a sharp decline of almost 32% fewer tourists than the previous year. However, the growth in the direct employment contibuted by travel and tourism industry maintained a steady growth. In this report, we will try to analyze the times series Nepal Tourism data starting from the year 2002 up till 2015. The variables used for analysis are:

1. TFXR - Total Foreign Exchange Reserves

2. TR - Number of Tourist Arrival

3. GFXE - Gross Foreign Exchange Earnings From Tourism

4. ALS - Average Length of Stay

5. TDCE - Travel & Tourism Direct Contribution to Employment


# Chapter 2: Data Analysis

For data analysis part we will do following trend analyses:

* __Trend line__

    $\widehat{Y} = b_0 + b_1x$  
      where   
      
      $\widehat{Y}$ = estimate of the dependent variable
      
      $x$ = coded value of the time variable
      
      $b_1$ = slope = $\frac{\sum xY}{\sum x^2}$
      
      $b_0$ = intercept = $\frac{\sum Y}{n}$
      

* __Percentages of trend__

    -  __Percent of Trend__
        
          We have
        
          $Percent of Trend$ = $\frac{Y}{\widehat{Y}} \times 100$ 
          
          where
          
          $Y$ = actual time-series value
          
          $\widehat{Y}$ = estimated trend value from the same point in the time series
          
          baseline = 100
        
    -  __Relative Cyclical Residual__
        
          We have
        
          $Relative Cyclical Residual$ = $\frac{Y - \widehat{Y}}{\widehat{Y}} \times 100$
        
          baseline = 0

* __Forecast__ value for the next three years, i.e. 2016, 2017, 2018

As our series contains data from 2002 to 2014, $n = 14$ 

## Transfering $X$ into $x$

- Since $n$ is even, $x = 2(X - \overline{X})$

# 2.1: Trend analysis of Total Foreign Exchange Reserves(TFXR)


```{r, echo=FALSE}

n <- nrow(nepal_tourism)
nsum <- n*(n+1)/2
x_bar <- nsum/n
row_vec <- row(nepal_tourism)[,1]
x <- 2*(row_vec-x_bar)
xY <- x * nepal_tourism[,2]
tfxr <- cbind(row_vec, nepal_tourism[,1:2], x, xY, x^2)
tfxr <- rbind(tfxr, colSums(tfxr))
#xtable(tfxr)
kable(tfxr, col.names = c("SN.", "Year ($X$)", "TFXR ($Y$)", "$x = 2(X - \\overline{X})$", "$xY$", "$x^2$"))

```


```{r, fig.width=5, fig.height=3, fig.cap="Data plot for Total Foreign Exchange Reserves augmented with LOESS regression curve with 95% confidence interval range", results='asis', echo=FALSE}

ggplot(tfxr[-15,], aes(x=Year, y=`Total Foreign Exchange Reserves`)) + geom_point() + geom_line(aes(y=`Total Foreign Exchange Reserves`)) + xlab("Year") + ylab("Total Foreign Exchange Reserves \n (in Billions Rs.)") + theme_light() + theme(legend.position="top", text = element_text(size=8)) + scale_y_continuous(labels = function(x)x/1000) + stat_smooth()
```

## Trend Line

* **Linear Trend**
  
    We have
      
    $\widehat{Y} = b_0 + b_1X$  
      where   
    ```{r, echo=FALSE}
    b_1 <- tfxr[15,5]/tfxr[15,6]
    b_0 <- tfxr[15,3]/n
    ```
      $b_1 = \frac{\sum xY}{\sum x^2}$ = $\frac{`r tfxr[15,5]`}{`r tfxr[15,6]`}$ = `r b_1`
      
      $b_0 = \frac{\sum Y}{n}$ = $\frac{`r tfxr[15,3]`}{`r n`}$ = `r b_0`  
      
      
    So, $\widehat{Y} = b_0 + b_1X = `r b_0` + `r b_1`x$
    
```{r, fig.width=5, fig.height=3, fig.cap="Linear trend line of Total Foreign Exchange Reserves with 95% confidence interval range" , results='asis', echo=FALSE}

ggplot(tfxr[-15,], aes(x=Year, y=`Total Foreign Exchange Reserves`)) + geom_point() + geom_line(aes(y=`Total Foreign Exchange Reserves`), color="#555555") + xlab("Year") + ylab("Total Foreign Exchange Reserves \n (in Billions Rs.)") + theme_light() + theme(legend.position="right", text = element_text(size=8)) + scale_y_continuous(labels = function(x)x/1000)  + stat_function(fun = function(x) {b_0 + b_1 * 2 * (x - 2001 - x_bar)}, size=1) + geom_smooth(method = "lm", color="black")
```

  
## Cyclical Variation 

*  __Percent of Trend__
    
    We have
    
    $Percent of Trend$ = $\frac{Y}{\widehat{Y}} \times 100$, where
    
    $Y$ = actual time-series value
    
    $\widehat{Y}$ = estimated trend value from the same point in the time series
    
    baseline = 100
    
*  __Relative Cyclical Residual__
    
    We have
    
    $Relative Cyclical Residual$ = $\frac{Y - \widehat{Y}}{\widehat{Y}} \times 100$, where
    
    baseline = 0
    
```{r, echo=FALSE}
y_hat <- b_0 + b_1 * tfxr[1:14, 4]
perTrend <- tfxr[1:14,3] / y_hat * 100
relCyc <- (tfxr[1:14,3] - y_hat) / y_hat * 100
tfxr <- cbind(tfxr[1:14, c(2, 3, 4)], y_hat, perTrend, relCyc)
kable(tfxr, col.names = c("Year ($X$)", "TFXR ($Y$)", "$x$", "$\\widehat{Y}$", "$\\frac{Y}{\\widehat{Y}} \\times 100$", "$\\frac{Y - \\widehat{Y}}{\\widehat{Y}} \\times 100$"))

```

  
For both measure of variaton, the highest fluctuation occured in the year 2002 where the *percent of trend* indicated the actual total foreign exchange reserves was about 878% of the expected value that year. For the same year, the *relative cyclical residual* indicated that the actual value was about 978% excess of the expected value.

  
  
```{r, fig.width=5, fig.height=3, fig.cap="Percent of trend around the trend line for Total Foreign Exchange Reserves", results='asis', echo=FALSE}

ggplot(tfxr[-15,], aes(x=Year, y=perTrend)) + geom_point() + geom_line(aes(y=`perTrend`)) + xlab("Year") + ylab("Percent of trend") + theme_light() + theme(legend.position="top", text = element_text(size=8)) + stat_function(fun = function(x) {x = 100}, linetype="dashed") + scale_y_continuous(breaks=seq(-900, 300, by=100))
```


## Forecasting Total Foreign Exchange Reserves for the years 2016, 2017, 2018

*  __2016__  
    ```{r, echo=FALSE}
    sn <- 15
    x <- 2*(sn - x_bar)
    p2016 <- b_0 + b_1 * x
    ```
    Here $X = `r sn`$,  
    $\therefore x = 2(`r x` - `r x_bar`)$ = `r x`  
    So, $\widehat{Y} = b_0 + b_1x = `r b_0` + `r b_1` * `r x`$ = $Rs.$ `r b_0 + b_1 * x` $million$

*  __2017__  
    ```{r, echo=FALSE}
    sn <- 16
    x <- 2*(sn - x_bar)
    p2017 <- b_0 + b_1 * x
    ```
    Here $X = `r sn`$,  
    $\therefore x = 2(`r x` - `r x_bar`)$ = `r x`  
    So, $\widehat{Y} = b_0 + b_1x = `r b_0` + `r b_1` * `r x`$ = $Rs.$ `r b_0 + b_1 * x` $million$

*  __2018__  
    ```{r, echo=FALSE}
    sn <- 17
    x <- 2*(sn - x_bar)
    p2018 <- b_0 + b_1 * x
    ```
    Here $X = `r sn`$,  
    $\therefore x = 2(`r x` - `r x_bar`)$ = `r x`  
    So, $\widehat{Y} = b_0 + b_1x = `r b_0` + `r b_1` * `r x`$ = $Rs.$ `r b_0 + b_1 * x` $million$



```{r, fig.width=5, fig.height=3, fig.cap="Forecasting of Total Foreign Exchange Reserves for the year 2016, 2017, 2018 using the linear trend line" , results='asis', echo=FALSE}

prediction <- data.frame(cbind(c(2016,2017,2018), c(p2016, p2017, p2018)))
colnames(prediction) <- c('Year', 'Total Foreign Exchange Reserves')

ggplot(tfxr[-15,], aes(x=Year, y=`Total Foreign Exchange Reserves`)) + geom_point(shape=23) + geom_line(aes(y=`Total Foreign Exchange Reserves`), color="#555555") + xlab("Year") + ylab("Total Foreign Exchange Reserves \n (in Billions Rs.)") + theme_light() + theme(legend.position="top", text = element_text(size=8)) + scale_y_continuous(labels = function(x)x/1000)  + stat_function(fun = function(x) {b_0 + b_1 * 2 * (x - 2001 - x_bar)}, size=1) + geom_smooth(method = "lm", color="black", se=FALSE) + geom_point(data = prediction, size=2, aes(shape = factor(Year, labels = c(paste("2016 : Rs.", round(p2016, digits=2), "million"), paste("2017 : Rs.", round(p2017, digits=2), "million"), paste("2018 : Rs.", round(p2018, digits=2), "million"))))) + labs(shape="")
```

# 2.2: Trend analysis of Tourist Arrival


```{r, echo=FALSE}

x <- 2*(row_vec-x_bar)
xY <- x * nepal_tourism[,3]
tfxr <- cbind(row_vec, nepal_tourism[,c(1,3)], x, xY, x^2)
tfxr <- rbind(tfxr, colSums(tfxr))
#xtable(tfxr)
kable(tfxr, col.names = c("SN.", "Year ($X$)", "Tourist Arrival ($Y$)", "$x = 2(X - \\overline{X})$", "$xY$", "$x^2$"))

```


```{r, fig.width=5, fig.height=3, fig.cap="Tourist Arrival augmented with LOESS regression curve with 95% confidence interval range", results='asis', echo=FALSE}

ggplot(tfxr[-15,], aes(x=Year, y=`Tourist Arrival`)) + geom_point() + geom_line(aes(y=`Tourist Arrival`)) + xlab("Year") + ylab("Tourist Arrival \n (in Thousands)") + theme_light() + theme(legend.position="top", text = element_text(size=8)) + stat_smooth() + scale_y_continuous(labels = function(x)x/1000)
```

## Trend Line

* **Linear Trend**
  
    We have
      
    $\widehat{Y} = b_0 + b_1X$  
      where   
    ```{r, echo=FALSE}
    b_1 <- tfxr[15,5]/tfxr[15,6]
    b_0 <- tfxr[15,3]/n
    ```
      $b_1 = \frac{\sum xY}{\sum x^2}$ = $\frac{`r tfxr[15,5]`}{`r tfxr[15,6]`}$ = `r b_1`
      
      $b_0 = \frac{\sum Y}{n}$ = $\frac{`r tfxr[15,3]`}{`r n`}$ = `r b_0`  
      
      
    So, $\widehat{Y} = b_0 + b_1X = `r b_0` + `r b_1`x$
    
```{r, fig.width=5, fig.height=3, fig.cap="Linear trend line of Tourist Arrival with 95% confidence interval range" , results='asis', echo=FALSE}

ggplot(tfxr[-15,], aes(x=Year, y=`Tourist Arrival`)) + geom_point() + geom_line(aes(y=`Tourist Arrival`), color="#555555") + xlab("Year") + ylab("Tourist Arrival \n (in Thousands)") + theme_light() + theme(legend.position="top", text = element_text(size=8))  + stat_function(fun = function(x) {b_0 + b_1 * 2 * (x - 2001 - x_bar)}, size=1) + scale_y_continuous(labels = function(x)x/1000) + geom_smooth(method = "lm", color="black")
```
  

## Cyclical Variation 

*  __Percent of Trend__
    
    We have
    
    $Percent of Trend$ = $\frac{Y}{\widehat{Y}} \times 100$, where
    
    $Y$ = actual time-series value
    
    $\widehat{Y}$ = estimated trend value from the same point in the time series
    
    baseline = 100
    
*  __Relative Cyclical Residual__
    
    We have
    
    $Relative Cyclical Residual$ = $\frac{Y - \widehat{Y}}{\widehat{Y}} \times 100$, where
    
    baseline = 0
    
```{r, echo=FALSE}
y_hat <- b_0 + b_1 * tfxr[1:14, 4]
perTrend <- tfxr[1:14,3] / y_hat * 100
relCyc <- (tfxr[1:14,3] - y_hat) / y_hat * 100
tfxr <- cbind(tfxr[1:14, c(2, 3, 4)], y_hat, perTrend, relCyc)
kable(tfxr, col.names = c("Year ($X$)", "Tourist Arrival ($Y$)", "$x$", "$\\widehat{Y}$", "$\\frac{Y}{\\widehat{Y}} \\times 100$", "$\\frac{Y - \\widehat{Y}}{\\widehat{Y}} \\times 100$"))

```

  
For both measure of variaton, the highest fluctuation occured in the year 2015 where the *percent of trend* indicated the actual total tourist arrival was about 69% of the expected value that year. For the same year, the *relative cyclical residual* indicated that the actual value was about 31% short of the expected value.

```{r, fig.width=5, fig.height=3, fig.cap="Percent of trend around the trend line for Tourist Arrival", results='asis', echo=FALSE}

ggplot(tfxr[-15,], aes(x=Year, y=perTrend)) + geom_point() + geom_line(aes(y=`perTrend`)) + xlab("Year") + ylab("Percent of trend") + theme_light() + theme(legend.position="top", text = element_text(size=8)) + stat_function(fun = function(x) {x = 100}, linetype="dashed") + scale_y_continuous(breaks=seq(60, 130, by=10))
```


  
## Forecasting Tourist Arrival for the years 2016, 2017, 2018

*  __2016__  
    ```{r, echo=FALSE}
    sn <- 15
    x <- 2*(sn - x_bar)
    p2016 <- round(b_0 + b_1 * x)
    ```
    Here $X = `r sn`$,  
    $\therefore x = 2(`r x` - `r x_bar`)$ = `r x`  
    So, $\widehat{Y} = b_0 + b_1x = `r b_0` + `r b_1` * `r x`$ $\simeq$ `r round(b_0 + b_1 * x)`

*  __2017__  
    ```{r, echo=FALSE}
    sn <- 16
    x <- 2*(sn - x_bar)
    p2017 <- round(b_0 + b_1 * x)
    ```
    Here $X = `r sn`$,  
    $\therefore x = 2(`r x` - `r x_bar`)$ = `r x`  
    So, $\widehat{Y} = b_0 + b_1x = `r b_0` + `r b_1` * `r x`$ $\simeq$ `r round(b_0 + b_1 * x)`

*  __2018__  
    ```{r, echo=FALSE}
    sn <- 17
    x <- 2*(sn - x_bar)
    p2018 <- round(b_0 + b_1 * x)
    ```
    Here $X = `r sn`$,  
    $\therefore x = 2(`r x` - `r x_bar`)$ = `r x`  
    So, $\widehat{Y} = b_0 + b_1x = `r b_0` + `r b_1` * `r x`$ $\simeq$ `r round(b_0 + b_1 * x)`



```{r, fig.width=5, fig.height=3, fig.cap="Forecasting of Tourist Arrival for the years 2016, 2017, 2018 using the linear trend line" , results='asis', echo=FALSE}

prediction <- data.frame(cbind(c(2016,2017,2018), c(p2016, p2017, p2018)))
colnames(prediction) <- c('Year', 'Tourist Arrival')

ggplot(tfxr[-15,], aes(x=Year, y=`Tourist Arrival`)) + geom_point(shape=23) + geom_line(aes(y=`Tourist Arrival`), color="#555555") + xlab("Year") + ylab("Tourist Arrival \n (in Thousands)") + theme_light() + theme(legend.position="top", text = element_text(size=8))  + stat_function(fun = function(x) {b_0 + b_1 * 2 * (x - 2001 - x_bar)}, size=1) + scale_y_continuous(labels = function(x)x/1000) + geom_smooth(method = "lm", color="black", se=FALSE) + geom_point(data = prediction, size=2, aes(shape = factor(Year, labels = c(paste("2016 : ", p2016), paste("2017 : ", p2017), paste("2018 : ", p2018))))) + labs(shape="")
```


# 2.3: Trend analysis of Gross Foreign Exchange Earnings From Tourism(GFXE)


```{r, echo=FALSE}

x <- 2*(row_vec-x_bar)
xY <- x * nepal_tourism[,4]
tfxr <- cbind(row_vec, nepal_tourism[,c(1,4)], x, xY, x^2)
tfxr <- rbind(tfxr, colSums(tfxr))
#xtable(tfxr)
kable(tfxr, col.names = c("SN.", "Year ($X$)", "GFXE ($Y$)", "$x = 2(X - \\overline{X})$", "$xY$", "$x^2$"), digits = 2)

```


```{r , fig.width=5, fig.height=3, fig.cap="Gross Foreign Exchange Earnings From Tourism augmented with LOESS regression curve with 95% confidence interval range", results='asis', echo=FALSE}

ggplot(tfxr[-15,], aes(x=Year, y=`Gross Foreign Exchange Earnings From Tourism`)) + geom_point() + geom_line(aes(y=`Gross Foreign Exchange Earnings From Tourism`)) + xlab("Year") + ylab("Gross Foreign Exchange Earnings From Tourism \n (in Billions Rs.)") + theme_light() + theme(legend.position="top", text = element_text(size=8)) + scale_y_continuous(labels = function(x)x/1000) + stat_smooth()
```

## Trend Line

* **Linear Trend**
  
    We have
      
    $\widehat{Y} = b_0 + b_1X$  
      where   
    ```{r, echo=FALSE}
        b_1 <- tfxr[15,5]/tfxr[15,6]
        b_0 <- tfxr[15,3]/n
    ```
      $b_1 = \frac{\sum xY}{\sum x^2}$ = $\frac{`r tfxr[15,5]`}{`r tfxr[15,6]`}$ = `r b_1`
      
      $b_0 = \frac{\sum Y}{n}$ = $\frac{`r tfxr[15,3]`}{`r n`}$ = `r b_0`  
      
      
    So, $\widehat{Y} = b_0 + b_1X = `r b_0` + `r b_1`x$
  
```{r , fig.width=5, fig.height=3, fig.cap="Linear trend line of Gross Foreign Exchange Earnings From Tourism with 95% confidence interval range" , results='asis', echo=FALSE}

ggplot(tfxr[-15,], aes(x=Year, y=`Gross Foreign Exchange Earnings From Tourism`)) + geom_point() + geom_line(aes(y=`Gross Foreign Exchange Earnings From Tourism`), color="#555555") + xlab("Year") + ylab("Gross Foreign Exchange Earnings From Tourism \n (in Billions Rs.)") + theme_light() + theme(legend.position="top", text = element_text(size=8)) + scale_y_continuous(labels = function(x)x/1000)  + stat_function(fun = function(x) {b_0 + b_1 * 2 * (x - 2001 - x_bar)}, size=1) + geom_smooth(method = "lm", color="black")
```

## Cyclical Variation 

*  __Percent of Trend__
    
    We have
    
    $Percent of Trend$ = $\frac{Y}{\widehat{Y}} \times 100$, where
    
    $Y$ = actual time-series value
    
    $\widehat{Y}$ = estimated trend value from the same point in the time series
    
    baseline = 100
    
*  __Relative Cyclical Residual__
    
    We have
    
    $Relative Cyclical Residual$ = $\frac{Y - \widehat{Y}}{\widehat{Y}} \times 100$, where
    
    baseline = 0
    
```{r, echo=FALSE}
y_hat <- b_0 + b_1 * tfxr[1:14, 4]
perTrend <- tfxr[1:14,3] / y_hat * 100
relCyc <- (tfxr[1:14,3] - y_hat) / y_hat * 100
tfxr <- cbind(tfxr[1:14, c(2, 3, 4)], y_hat, perTrend, relCyc)
kable(tfxr, col.names = c("Year ($X$)", "GFXE ($Y$)", "$x$", "$\\widehat{Y}$", "$\\frac{Y}{\\widehat{Y}} \\times 100$", "$\\frac{Y - \\widehat{Y}}{\\widehat{Y}} \\times 100$"))

```

  
For both measure of variaton, the highest fluctuation occured in the year 2002 where the *percent of trend* indicated the actual Gross Foreign Exchange Earnings From Tourism was about 215% of the expected value that year. For the same year, the *relative cyclical residual* indicated that the actual value was about 115% excess of the expected value.

```{r, fig.width=5, fig.height=3, fig.cap="Percent of trend around the trend line for Gross Foreign Exchange Earnings From Tourism", results='asis', echo=FALSE}

ggplot(tfxr[-15,], aes(x=Year, y=perTrend)) + geom_point() + geom_line(aes(y=`perTrend`)) + xlab("Year") + ylab("Percent of trend") + theme_light() + theme(legend.position="top", text = element_text(size=8)) + stat_function(fun = function(x) {x = 100}, linetype="dashed") + scale_y_continuous(breaks=seq(60, 220, by=20))
```


  
## Forecasting Gross Foreign Exchange Earnings From Tourism for the years 2016, 2017, 2018

*  __2016__  
    ```{r, echo=FALSE}
    sn <- 15
    x <- 2*(sn - x_bar)
    p2016 <- b_0 + b_1 * x
    ```
    Here $X = `r sn`$,  
    $\therefore x = 2(`r x` - `r x_bar`)$ = `r x`  
    So, $\widehat{Y} = b_0 + b_1x = `r b_0` + `r b_1` * `r x`$ = $Rs.$ `r b_0 + b_1 * x` $million$

*  __2017__  
    ```{r, echo=FALSE}
    sn <- 16
    x <- 2*(sn - x_bar)
    p2017 <- b_0 + b_1 * x
    ```
    Here $X = `r sn`$,  
    $\therefore x = 2(`r x` - `r x_bar`)$ = `r x`  
    So, $\widehat{Y} = b_0 + b_1x = `r b_0` + `r b_1` * `r x`$ = $Rs.$ `r b_0 + b_1 * x` $million$

*  __2018__  
    ```{r, echo=FALSE}
    sn <- 17
    x <- 2*(sn - x_bar)
    p2018 <- b_0 + b_1 * x
    ```
    Here $X = `r sn`$,  
    $\therefore x = 2(`r x` - `r x_bar`)$ = `r x`  
    So, $\widehat{Y} = b_0 + b_1x = `r b_0` + `r b_1` * `r x`$ = $Rs.$ `r b_0 + b_1 * x` $million$


```{r, fig.width=5, fig.height=3, fig.cap="Forecasting of Gross Foreign Exchange Earnings From Tourism for the years 2016, 2017, 2018 using the linear trend line" , results='asis', echo=FALSE}

prediction <- data.frame(cbind(c(2016,2017,2018), c(p2016, p2017, p2018)))
colnames(prediction) <- c('Year', 'Gross Foreign Exchange Earnings From Tourism')

ggplot(tfxr[-15,], aes(x=Year, y=`Gross Foreign Exchange Earnings From Tourism`)) + geom_point(shape=23) + geom_line(aes(y=`Gross Foreign Exchange Earnings From Tourism`), color="#555555") + xlab("Year") + ylab("Gross Foreign Exchange Earnings From Tourism \n (in Billions Rs.)") + theme_light() + theme(legend.position="top", text = element_text(size=8)) + scale_y_continuous(labels = function(x)x/1000)  + stat_function(fun = function(x) {b_0 + b_1 * 2 * (x - 2001 - x_bar)}, size=1) + geom_smooth(method = "lm", color="black", se=FALSE) + geom_point(data = prediction, size=2, aes(shape = factor(Year, labels = c(paste("2016 : Rs.", round(p2016, digits=2), "million"), paste("2017 : Rs.", round(p2017, digits=2), "million"), paste("2018 : Rs.", round(p2018, digits=2), "million"))))) + labs(shape="")
```



# 2.4: Average Length of Stay(ALS)


```{r, echo=FALSE}

x <- 2*(row_vec-x_bar)
xY <- x * nepal_tourism[,5]
tfxr <- cbind(row_vec, nepal_tourism[,c(1,5)], x, xY, x^2)
tfxr <- rbind(tfxr, colSums(tfxr))
#xtable(tfxr)
kable(tfxr, col.names = c("SN.", "Year ($X$)", "ALS ($Y$)", "$x = 2(X - \\overline{X})$", "$xY$", "$x^2$"))

```


```{r, fig.width=5, fig.height=3, fig.cap="Average Length of Stay augmented with LOESS regression curve with 95% confidence interval range", results='asis', echo=FALSE}

ggplot(tfxr[-15,], aes(x=Year, y=`Average Length of Stay`)) + geom_point() + geom_line(aes(y=`Average Length of Stay`)) + xlab("Year") + ylab("Average Length of Stay \n (in Days)") + theme_light() + theme(legend.position="top", text = element_text(size=8)) + stat_smooth()
```

## Trend Line

* **Linear Trend**
  
    We have
      
    $\widehat{Y} = b_0 + b_1X$  
      where   
    ```{r, echo=FALSE}
    b_1 <- tfxr[15,5]/tfxr[15,6]
    b_0 <- tfxr[15,3]/n
    ```
      $b_1 = \frac{\sum xY}{\sum x^2}$ = $\frac{`r tfxr[15,5]`}{`r tfxr[15,6]`}$ = `r b_1`
      
      $b_0 = \frac{\sum Y}{n}$ = $\frac{`r tfxr[15,3]`}{`r n`}$ = `r b_0`  
      
      
    So, $\widehat{Y} = b_0 + b_1X = `r b_0` + `r b_1`x$
  
```{r, fig.width=5, fig.height=3, fig.cap="Linear trend line of Average Length of Stay with 95% confidence interval range" , results='asis', echo=FALSE}

ggplot(tfxr[-15,], aes(x=Year, y=`Average Length of Stay`)) + geom_point() + geom_line(aes(y=`Average Length of Stay`), color="#555555") + xlab("Year") + ylab("Average Length of Stay \n (in Days)") + theme_light() + theme(legend.position="top", text = element_text(size=8)) + stat_function(fun = function(x) {b_0 + b_1 * 2 * (x - 2001 - x_bar)}, size=1) + geom_smooth(method = "lm", color="black")
```

## Cyclical Variation 

*  __Percent of Trend__
    
    We have
    
    $Percent of Trend$ = $\frac{Y}{\widehat{Y}} \times 100$, where
    
    $Y$ = actual time-series value
    
    $\widehat{Y}$ = estimated trend value from the same point in the time series
    
    baseline = 100
    
*  __Relative Cyclical Residual__
    
    We have
    
    $Relative Cyclical Residual$ = $\frac{Y - \widehat{Y}}{\widehat{Y}} \times 100$, where
    
    baseline = 0
    
```{r, echo=FALSE}
y_hat <- b_0 + b_1 * tfxr[1:14, 4]
perTrend <- tfxr[1:14,3] / y_hat * 100
relCyc <- (tfxr[1:14,3] - y_hat) / y_hat * 100
tfxr <- cbind(tfxr[1:14, c(2, 3, 4)], y_hat, perTrend, relCyc)
kable(tfxr, col.names = c("Year ($X$)", "ALS ($Y$)", "$x$", "$\\widehat{Y}$", "$\\frac{Y}{\\widehat{Y}} \\times 100$", "$\\frac{Y - \\widehat{Y}}{\\widehat{Y}} \\times 100$"))

```

  
For both measure of variaton, the highest fluctuation occured in the year 2004 where the *percent of trend* indicated the actual Average Length of Stay was about 132% of the expected value that year. For the same year, the *relative cyclical residual* indicated that the actual value was about 32% excess of the expected value.

```{r, fig.width=5, fig.height=3, fig.cap="Percent of trend around the trend line for Average Length of Stay", results='asis', echo=FALSE}

ggplot(tfxr[-15,], aes(x=Year, y=perTrend)) + geom_point() + geom_line(aes(y=`perTrend`)) + xlab("Year") + ylab("Percent of trend") + theme_light() + theme(legend.position="top", text = element_text(size=8)) + stat_function(fun = function(x) {x = 100}, linetype="dashed") + scale_y_continuous(breaks=seq(80, 140, by=20))
```

 
  
## Forecasting Average Length of Stay for the years 2016, 2017, 2018

*  __2016__  
    ```{r, echo=FALSE}
    sn <- 15
    x <- 2*(sn - x_bar)
    p2016 <- round(b_0 + b_1 * x, 2)
    ```
    Here $X = `r sn`$,  
    $\therefore x = 2(`r x` - `r x_bar`)$ = `r x`  
    So, $\widehat{Y} = b_0 + b_1x = `r b_0` + `r b_1` * `r x`$ $\simeq$ `r b_0 + b_1 * x` $days$

*  __2017__  
    ```{r, echo=FALSE}
    sn <- 16
    x <- 2*(sn - x_bar)
    p2017 <- round(b_0 + b_1 * x, 2)
    ```
    Here $X = `r sn`$,  
    $\therefore x = 2(`r x` - `r x_bar`)$ = `r x`  
    So, $\widehat{Y} = b_0 + b_1x = `r b_0` + `r b_1` * `r x`$ $\simeq$ `r b_0 + b_1 * x` $days$

*  __2018__  
    ```{r, echo=FALSE}
    sn <- 17
    x <- 2*(sn - x_bar)
    p2018 <- round(b_0 + b_1 * x, 2)
    ```
    Here $X = `r sn`$,  
    $\therefore x = 2(`r x` - `r x_bar`)$ = `r x`  
    So, $\widehat{Y} = b_0 + b_1x = `r b_0` + `r b_1` * `r x`$ $\simeq$ `r b_0 + b_1 * x` $days$

```{r, fig.width=5, fig.height=3, fig.cap="Forecasting of Average Length of Stay for the years 2016, 2017, 2018 using the linear trend line" , results='asis', echo=FALSE}

prediction <- data.frame(cbind(c(2016,2017,2018), c(p2016, p2017, p2018)))
colnames(prediction) <- c('Year', 'Average Length of Stay')

ggplot(tfxr[-15,], aes(x=Year, y=`Average Length of Stay`)) + geom_point(shape=23) + geom_line(aes(y=`Average Length of Stay`), color="#555555") + xlab("Year") + ylab("Average Length of Stay \n (in Days)") + theme_light() + theme(legend.position="top", text = element_text(size=8))  + stat_function(fun = function(x) {b_0 + b_1 * 2 * (x - 2001 - x_bar)}, size=1) + scale_y_continuous(labels = function(x)x/1000) + geom_smooth(method = "lm", color="black", se=FALSE) + geom_point(data = prediction, size=2, aes(shape = factor(Year, labels = c(paste("2016 : ", p2016, "days "), paste("2017 : ", p2017, "days "), paste("2018 : ", p2018, "days "))))) + labs(shape="")
```


# 2.5: Travel & Tourism Direct Contribution to Employment(TDCE)


```{r, echo=FALSE}

x <- 2*(row_vec-x_bar)
xY <- x * nepal_tourism[,6]
tfxr <- cbind(row_vec, nepal_tourism[,c(1,6)], x, xY, x^2)
tfxr <- rbind(tfxr, colSums(tfxr))
#xtable(tfxr)
kable(tfxr, col.names = c("SN.", "Year ($X$)", "TDCE ($Y$)", "$x = 2(X - \\overline{X})$", "$xY$", "$x^2$"))

```


```{r, fig.width=5, fig.height=3, fig.cap="Travel & Tourism Direct Contribution to Employment augmented with LOESS regression curve with 95% confidence interval range", results='asis', echo=FALSE}

ggplot(tfxr[-15,], aes(x=Year, y=`Travel & Tourism Direct Contribution to Employment`)) + geom_point() + geom_line(aes(y=`Travel & Tourism Direct Contribution to Employment`)) + xlab("Year") + ylab("Travel & Tourism Direct Contribution to Employment \n (in Thousands)") + theme_light() + theme(legend.position="top", text = element_text(size=8)) + stat_smooth()  + scale_y_continuous(labels = function(x)x/1000)
```

## Trend Line

* **Linear Trend**
  
    We have
      
    $\widehat{Y} = b_0 + b_1X$  
      where   
    ```{r, echo=FALSE}
    b_1 <- tfxr[15,5]/tfxr[15,6]
    b_0 <- tfxr[15,3]/n
    ```
      $b_1 = \frac{\sum xY}{\sum x^2}$ = $\frac{`r tfxr[15,5]`}{`r tfxr[15,6]`}$ = `r b_1`
      
      $b_0 = \frac{\sum Y}{n}$ = $\frac{`r tfxr[15,3]`}{`r n`}$ = `r b_0`  
      
      
    So, $\widehat{Y} = b_0 + b_1X = `r b_0` + `r b_1`x$
  
```{r, fig.width=5, fig.height=3, fig.cap="Linear trend line of Travel & Tourism Direct Contribution to Employment with 95% confidence interval range" , results='asis', echo=FALSE}

ggplot(tfxr[-15,], aes(x=Year, y=`Travel & Tourism Direct Contribution to Employment`)) + geom_point() + geom_line(aes(y=`Travel & Tourism Direct Contribution to Employment`), color="#555555") + xlab("Year") + ylab("Travel & Tourism Direct Contribution to Employment \n (in Thousands)") + theme_light() + theme(legend.position="top", text = element_text(size=8)) + stat_function(fun = function(x) {b_0 + b_1 * 2 * (x - 2001 - x_bar)}, size=1) + scale_y_continuous(labels = function(x)x/1000) + geom_smooth(method = "lm", color="black")
```

## Cyclical Variation 

*  __Percent of Trend__
    
    We have
    
    $Percent of Trend$ = $\frac{Y}{\widehat{Y}} \times 100$, where
    
    $Y$ = actual time-series value
    
    $\widehat{Y}$ = estimated trend value from the same point in the time series
    
    baseline = 100
    
*  __Relative Cyclical Residual__
    
    We have
    
    $Relative Cyclical Residual$ = $\frac{Y - \widehat{Y}}{\widehat{Y}} \times 100$, where
    
    baseline = 0
    
```{r, echo=FALSE}
y_hat <- b_0 + b_1 * tfxr[1:14, 4]
perTrend <- tfxr[1:14,3] / y_hat * 100
relCyc <- (tfxr[1:14,3] - y_hat) / y_hat * 100
tfxr <- cbind(tfxr[1:14, c(2, 3, 4)], y_hat, perTrend, relCyc)
kable(tfxr, col.names = c("Year ($X$)", "TDCE ($Y$)", "$x$", "$\\widehat{Y}$", "$\\frac{Y}{\\widehat{Y}} \\times 100$", "$\\frac{Y - \\widehat{Y}}{\\widehat{Y}} \\times 100$"))

```

  
For both measure of variaton, the highest fluctuation occured in the year 2004 where the *percent of trend* indicated the actual Travel & Tourism Direct Contribution to Employment was about 119% of the expected value that year. For the same year, the *relative cyclical residual* indicated that the actual value was about 19% excess of the expected value.

```{r, fig.width=5, fig.height=3, fig.cap="Percent of trend around the trend line for Travel & Tourism Direct Contribution to Employment", results='asis', echo=FALSE}

ggplot(tfxr[-15,], aes(x=Year, y=perTrend)) + geom_point() + geom_line(aes(y=`perTrend`)) + xlab("Year") + ylab("Percent of trend") + theme_light() + theme(legend.position="top", text = element_text(size=8)) + stat_function(fun = function(x) {x = 100}, linetype="dashed") + scale_y_continuous(breaks=seq(60, 140, by=20))
```

  
  
## Forecasting Travel & Tourism Direct Contribution to Employment for the years 2016, 2017, 2018

*  __2016__  
    ```{r, echo=FALSE}
    sn <- 15
    x <- 2*(sn - x_bar)
    p2016 <- round(b_0 + b_1 * x)
    ```
    Here $X = `r sn`$,  
    $\therefore x = 2(`r x` - `r x_bar`)$ = `r x`  
    So, $\widehat{Y} = b_0 + b_1x = `r b_0` + `r b_1` * `r x`$ $\simeq$ `r round(b_0 + b_1 * x)`

*  __2017__  
    ```{r, echo=FALSE}
    sn <- 16
    x <- 2*(sn - x_bar)
    p2017 <- round(b_0 + b_1 * x)
    ```
    Here $X = `r sn`$,  
    $\therefore x = 2(`r x` - `r x_bar`)$ = `r x`  
    So, $\widehat{Y} = b_0 + b_1x = `r b_0` + `r b_1` * `r x`$ $\simeq$ `r round(b_0 + b_1 * x)`

*  __2018__  
    ```{r, echo=FALSE}
    sn <- 17
    x <- 2*(sn - x_bar)
    p2018 <- round(b_0 + b_1 * x)
    ```
    Here $X = `r sn`$,  
    $\therefore x = 2(`r x` - `r x_bar`)$ = `r x`  
    So, $\widehat{Y} = b_0 + b_1x = `r b_0` + `r b_1` * `r x`$ $\simeq$ `r round(b_0 + b_1 * x)`

```{r, fig.width=5, fig.height=3, fig.cap="Forecasting of Travel & Tourism Direct Contribution to Employment for the years 2016, 2017, 2018 using the linear trend line" , results='asis', echo=FALSE}

prediction <- data.frame(cbind(c(2016,2017,2018), c(p2016, p2017, p2018)))
colnames(prediction) <- c('Year', 'Travel & Tourism Direct Contribution to Employment')

ggplot(tfxr[-15,], aes(x=Year, y=`Travel & Tourism Direct Contribution to Employment`)) + geom_point(shape=23) + geom_line(aes(y=`Travel & Tourism Direct Contribution to Employment`), color="#555555") + xlab("Year") + ylab("Travel & Tourism Direct Contribution to Employment \n (in Thousands)") + theme_light() + theme(legend.position="top", text = element_text(size=8))  + stat_function(fun = function(x) {b_0 + b_1 * 2 * (x - 2001 - x_bar)}, size=1) + scale_y_continuous(labels = function(x)x/1000) + geom_smooth(method = "lm", color="black", se=FALSE) + geom_point(data = prediction, size=2, aes(shape = factor(Year, labels = c(paste("2016 : ", p2016), paste("2017 : ", p2017), paste("2018 : ", p2018))))) + labs(shape="")
```


# Citation

Turner, R.(2015). *Travel & Tourism: Economic Impact 2015(Nepal).* Retrieved from World Travel & Tourism Council Website: http://sp.wttc.org/-/media/files/reports/economic-impact-research/countries-2015/nepal2015.pdf
  

Ministry of Culture, Tourism & Civil Aviation.(2016). *NEPAL TOURISM STATISTICS 2015.* Retrieved from Government of Nepal Ministry of Culture, Tourism & Civil Aviation Website: http://www.tourism.gov.np/new/downloadfile/Nepal_Tourism_Statistics_2015_forwebsite_edited.pdf

  
Nepal - Travel & Tourism Direct Contribution to Employment. (n.d.). Retrieved October 6, 2016, from https://knoema.com/atlas/Nepal/topics/Tourism/Travel-and-Tourism-Direct-Contribution-to-Employment/Direct-Contribution-to-Employment-thousands